---
title: "R Code for Analysis of Customer Churn Prediction"
author: '21203863'
date: "2022-07-31"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Loading packages required for Analysis:-**
```{r}
library(ggplot2)
library(cowplot)
library(stringr)
library(rpart)
library(partykit)
library(pROC)
library(caret)
```
**Import data:-**
```{r}
churn_data = read.csv("Churn Data.csv")
str(churn_data)
```
# Exploratory Data Analysis
**Check for null values:-**
```{r}
apply(X = is.na(churn_data), MARGIN = 2, FUN = sum)
```
**Remove Null values:-**
```{r}
churn_data<-na.omit(churn_data)
dim(churn_data)
```
**Analysis of Churn variable:-**
```{r}
ggplot(churn_data, aes(Churn)) + 
  geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn))) + 
  scale_y_continuous(labels=scales::percent) + theme_bw()+ xlab("Churn") + 
  ylab("Percent")+ ggtitle("Churn Percent")+ theme(legend.position="none")
```
**Analysis of customer demographics:-**
```{r}
plot_grid(

ggplot(churn_data, aes(x=gender,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)))+ scale_y_continuous(labels=scales::percent) + ylab("Percent") + theme_bw()+theme(legend.position="none"),
  
ggplot(churn_data, aes(x=SeniorCitizen,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none"),
  
ggplot(churn_data, aes(x=Partner,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw(),
  
ggplot(churn_data, aes(x=Dependents,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none"))
```
**Analysis of Service information:-**
```{r}
plot_grid(

ggplot(churn_data, aes(x=PhoneService,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none"),
         
ggplot(churn_data, aes(x=MultipleLines,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none") + 
          scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),

ggplot(churn_data, aes(x=InternetService,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none"), 

ggplot(churn_data, aes(x=OnlineSecurity,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none")+scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),

ggplot(churn_data, aes(x=OnlineBackup,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none")+ scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          
ggplot(churn_data, aes(x=DeviceProtection,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw()+theme(legend.position="none")+ scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          
ggplot(churn_data, aes(x=TechSupport,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw() +theme(legend.position="none")+ scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),
          
ggplot(churn_data, aes(x=StreamingTV,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw() +
theme(legend.position="none") + scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),

ggplot(churn_data, aes(x=StreamingMovies,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw() + scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
)
```
**Analysis of Customer Information:-**
*Categorical data:-*
```{r}
plot_grid(
ggplot(churn_data, aes(x=Contract,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw() + scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),

ggplot(churn_data, aes(x=PaperlessBilling,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw() + theme(legend.position="none") + scale_x_discrete(labels = function(x) str_wrap(x, width = 10)),

ggplot(churn_data, aes(x=PaymentMethod,fill=Churn))+ geom_bar(aes(y = (..count..)/sum(..count..), fill = factor(Churn)),position = 'fill')+scale_y_continuous(labels=scales::percent) + ylab("Percent")+theme_bw() + theme(legend.position="none") + scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

)
```
*Numerical data:-*
```{r}
plot_grid(
ggplot(churn_data, aes(y= tenure, x = "", fill = Churn)) + 
geom_boxplot()+ 
theme_bw()+ theme(legend.position="none")+
xlab(" "),

ggplot(churn_data, aes(y= MonthlyCharges, x = "", fill = Churn)) + 
geom_boxplot()+ 
theme_bw()+ theme(legend.position="none")+
xlab(" "),

ggplot(churn_data, aes(y= TotalCharges, x = "", fill = Churn)) + 
geom_boxplot()+ 
theme_bw()+
xlab(" ")
)
```
# Data Preprocessing
**Preprocessing steps:-**
```{r}
# Re-encode No internet Service and No as No
churn_data<- data.frame(lapply(churn_data, function(x) {
  gsub("No internet service", "No", x)}))
# Re-encode No phone service and No as No
churn_data<- data.frame(lapply(churn_data, function(x) {
  gsub("No phone service", "No", x)}))
# Remove numerical variables and scale them
num_columns <- c("tenure", "MonthlyCharges", "TotalCharges")
churn_data[num_columns] <- sapply(churn_data[num_columns], as.numeric)

churn_int <- churn_data[,c("tenure", "MonthlyCharges", "TotalCharges")]
churn_int <- data.frame(scale(churn_int))
# Remove categorical variables and create dummy variables
churn_cat <- churn_data[,-c(1,6,19,20)]

#Creating Dummy Variables
dummy<- data.frame(sapply(churn_cat,function(x) data.frame(model.matrix(~x-1,data =churn_cat))[,-1]))

#Combining the data
churn_final <- cbind(churn_int,dummy)
# Response variable as factor
churn_final$Churn <-as.factor(churn_final$Churn)
```
# Modelling
**Split data to train and test:-**
```{r}
set.seed(123)
indices = createDataPartition(churn_final$Churn, p = 0.70, list = FALSE)
train = churn_final[indices,]
validation = churn_final[-indices,]
```
**Fit Logistic Regression model and save the model:-**
```{r}
fit <- glm(Churn ~ ., data = train, family = "binomial")
save(fit, file = 'LogisticRegression.rda')
phat <- predict(fit,newdata=validation[,-24],type = "response")
tau <- 0.5
pred <- ifelse(phat > tau, 1, 0)
cm_lr<-confusionMatrix(as.factor(pred), as.factor(validation$Churn))
cm_lr
```
**Fit Decision Trees and save the model:-**
```{r}
grid <- 10^seq(10, -2, length = 100)
train_ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 10)
tune_grid <- expand.grid(cp = grid)
fit_dtr <- train(Churn ~ ., data = train,
                method = "rpart",
                trControl = train_ctrl,
                tuneGrid = tune_grid)
fit_dt=rpart(Churn~.,data=train,control=rpart.control(cp=fit_dtr$bestTune[[1]]))
save(fit_dt, file = 'DecisionTrees.rda')
predict_unseen <-predict(fit_dt,validation[,-24],type='class')
cm_dt<-confusionMatrix(validation$Churn,predict_unseen,positive="0")
cm_dt
plot(as.party(fit_dt))
```
#Model Evaluation
**ROC-AUC curve:-**
```{r}
glm.roc <- roc(response = validation$Churn, predictor = as.numeric(pred))
DT.roc <- roc(response = validation$Churn, predictor = as.numeric(predict_unseen))
plot(glm.roc,legacy.axes = TRUE, print.auc.y = 1.0, print.auc = TRUE)
plot(DT.roc, col = "blue", add = TRUE, print.auc.y = 0.65, print.auc = TRUE)
legend("bottomright", c("Decision Tree", "Logistic"),
       lty = c(1,1), lwd = c(2, 2), col = c("blue", "black"), cex = 0.75)
```